Further Explorations of \suldiox~Aqueous Adsorption:

### 2011-08-31 Cycle-Results.tex
About to start final edits to the pre-draft sans energy stuff

Knowing that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and also often taking part in some type of cyclic structure, how long does the cyclic hydrate form before breaking to an acyclic hydrate structure? To answer the questions of cyclic lifespans, a method was devised to define a lifetime of a cycle. For each trajectory, the coordinate data was analyzed to determine if a cyclic hydrate structure was formed based on the bonding distance criteria established earlier in this manuscript. A timeline was then produced where each timestep was given a value of 1 or 0 depending on if a \suldiox~hydrate bonding cycle was present or not, respectively. This resulted in a time-function, $C(t)$, similar in nature to a digital signal.

In an electro-mechanical system, a mechanical switch triggering an electrical signal often outputs a noisy signal that has a period of signal bounce before settling to a final value. This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing. Many mechanical, electrical, and software solutions have been devised for this problem of signal bounce. Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before breaking, where the $C(t)$ function bounces before settling into a final value. The bounce in the function, manifested as a series of rapid switches between cycle presence and absence, signals a transition period due to the distortions of the bonds in the cyclic hydrates. We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions. The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states. $f(t)$ may then be used to calculate statistics about the cycle lifespans. The statistics thus calculated will no longer be skewed by the multiple short changes of state associated with the function bounce.

A representative portion of the resulting cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time. The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the $C(t)$ function lasting less than 10 fs each. The following procedure was used to debounce the time-function to disregard very brief moments when the cycle breaks or forms for less than 20 fs. The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width. The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:

@@

where $f(t)$ is the debounced time-function that better represents the lifespans of cycles, eliminating the very brief discontinuities and excursions outside the bonding criteria that cause the rapid bouncing of the cyclic time-function, $C(t)$. Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).

@@

We know from the transition frequency plots of Figure \ref{fig:coordination-transitions} that the bonding coordinations are switching frequently. The \suldiox~is likely forming and breaking bonds with waters extraneous to the cyclic hydrate structures (i.e. not directly involved in the bonds of the cycle). For the longer-lived cycles, the extraneous bonding to the \suldiox~may have little effect on the cyclic hydrate waters. However, any time the \suldiox~switches into an unbound, sulfur-only, or oxygen-only coordination (i.e. ``S'', ``SS'', ``O'', etc), the cyclic structure is necessarily broken. Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure. Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively. Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold). Of the three most encountered bonding coordinations, the hot system shifts population from the ``S'' (a cycle-breaking coordination) to ``SO'' and ``SOO'' (coordinations that allow for bonding cycle formation through the \suldiox). This is relative to the cold system coordination distribution that has equal ``S'' and ``SOO'' populations. The increased temperature appears to cause the \suldiox~to bond in a way that is more conducive to the formation of cyclic structures. This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.

>>>>

We know that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and is also often taking part in some type of cyclic structure. With this in mind, it becomes interesting to ask: how long does a cyclic hydrate form before breaking to an acyclic hydrate structure? To answer the question of cyclic lifespan, a method was devised to define the lifetime of a cycle. For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript. A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a particular \suldiox~hydrate bonding cycle was present or not, respectively. This resulted in a time-function, $C(t)$, similar in nature to a time-varying digital signal.

In an electro-mechanical system, a mechanical switch often outputs a noisy signal, full of transients or ``signal bounce'' before settling to a final value. This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing. Many mechanical, electrical, and software solutions have been devised to suppress the transients or ``debounce'' the signal. Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before dissolution, where the $C(t)$ function bounces before settling into a final value. The bounce in the function, manifests itself in our statistics as a series of rapid switches between cycle presence and cycle absence. Physically, the bond lengths within the cycle (or the cycle-to-be) are fluctuating back and forth across our bond-length criteria. It is thus an artifact of our algorithmic determination of the presence or absence of a bond. Removing this artifact will allow us to examine the longer-time bonding behavior and this is accomplished as follows.

We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions. The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states. $f(t)$ may then be used to calculate statistics about the cycle lifespans. A representative portion of a cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time. The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each. The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration. The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:

@@

where $f(t)$ is the debounced time-function that represents the lifespans of cycles. Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).

@@

We know from the transition frequency plots of Figure \ref{fig:coordination-transitions} that the bonding coordinations are switching frequently. The \suldiox~is likely forming and breaking bonds with waters external to the cyclic hydrate structures (i.e. not directly involved in the bonds of the cycle). For the longer-lived cycles, the external bonding to the \suldiox~may have little effect on the cyclic hydrate waters. However, any time the \suldiox~switches into an unbound, sulfur-only, or oxygen-only coordination (i.e. ``S'', ``SS'', ``O'', etc), the cyclic structure is necessarily broken. Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure. Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively. Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold). Of the three most encountered bonding coordinations, the hot system shifts population from the ``S'' (a cycle-breaking coordination) to ``SO'' and ``SOO'' (coordinations that allow for bonding cycle formation through the \suldiox). This is relative to the cold system coordination distribution that has equal ``S'' and ``SOO'' populations. The increased temperature appears to cause the \suldiox~to bond in a way that is more conducive to the formation of cyclic structures. This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.

### 2011-08-31 Introduction.tex
About to start final edits to the pre-draft sans energy stuff

%The chemistry of our atmosphere is vastly complex and poorly understood. Reactions taking place on particle surfaces and liquid boundaries dominate many atmospheric chemical pathways, and guide terrestrial and industrial byproducts to their ultimate forms and destinations. How do atmospheric conditions affect the chemistry of environmental constituents?

>>>>

One of Chemistry's many remaining ``black box'' events is the adsorption of gas molecules onto a water surface. Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions. How does gas initially bind to a water surface, and what steps are involved in the subsequent adsorption? How does an unbound gas molecule near a water surface affect the water to which it will bind? What is the structure of hydrating waters in the surface region, and how does a hydrated solute molecule behave differently than as a gas? Experiments to address these questions provide valuable information, but can never fully characterize these microscopic events and behaviors. However, these systems can be fully characterized computationally, and when coupled to the previous experimental work can provide a much more complete picture of gaseous adsorption to aqueous surfaces.

As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems. Its molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions. A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in interfacial water regions.

In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on water. We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the variability of the \suldiox~hydrate structures, and the behaviors of the water surface. The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD. It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated. In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas. Understanding the orientational behavior of molecules in the aqueous interfacial region during adsorption was a necessary first step to understanding the specific details of gas binding and surface behavior.

%Unlike small cluster studies that use DFT calculations on geometry optimized structures to find likely geometries in a static vacuum setting, The quantum MD technique allows us to simulate the process much more realistically, and with greater accuracy than our previous classical MD work could allow.\cite{Shamay2011} This computational study greatly enhances the picture developed in our previous work on \suldiox.

Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water. Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 270K. This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Conclusions regarding the specific nature of those complexes were inferred from the experiments. With our current computational studies we now have insights about the specific microscopic geometries and behaviors of the hydrating complexes.

We believe this to be the first temperature study using quantum MD to study the binding of small gas molecules on a water surface. We show how temperature affects the bonding behavior of the surface adsorbed \suldiox~to neighboring waters, and propose a sequential binding mechanism for \suldiox~adsorbing to a water surface. We also examine \suldiox~behavior when bound to the surface waters. Lastly, our analysis demonstrates an extended bonding of \suldiox~hydrate structures, as \suldiox~hydrates are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.
